Digital History in Finland, 1488-1828

### 2016-07-18 inst/examples/20160715-Krakow-Fennica.Rmd
Rmd update

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed
df.full <- df.combined.preprocessed %>% filter(!remove)
# Data with years limited as well
df0 <- df.full %>% filter(publication_year >= min.year & publication_year <= max.year)
```

@@

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p1 <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none")

df <- df0 %>% filter(catalog == "Kungliga")
# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper.consumption.km2, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

@@

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0
df <- df %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

df <- df %>% filter(catalog == catalogue & publication_place == place)

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper.consumption.km2)

@@

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

  p <- ggplot(df2, aes(y = paper.consumption.km2,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(catal)
  print(p)
}
```

>>>>

```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Full combined catalogue (Fen + Kun) with marked duplicates
df.combined.preprocessed <- readRDS("df.combined.Rds")
# Data with duplicates removed and years limited
df0 <- df.combined.preprocessed %>% filter(!remove) %>%
	   			    filter(publication_year >= min.year &
					   publication_year <= max.year)
```

@@

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}

library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p1 <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none")

# Add indicator of the publication period
df <- df0 %>% filter(catalog == "Kungliga")
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>% group_by(period, paper)
df <- df %>% summarise(n = n(),
		  paper2 = sum(paper, na.rm = TRUE)) %>%
             mutate(paper.per.title = paper2/n)		 
df$paper <- df$paper2
df$paper2 <- NULL

@@

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
df <- df0 %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

@@

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue <- "Fennica"

df <- df0 %>% filter(catalog == catalogue & publication_place == place)

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper)

@@

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper, na.rm = TRUE))

@@

sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper, na.rm = TRUE))

@@

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- df0 %>% filter(catalog == catal) %>% 
                 group_by(publication_decade, gatherings) %>% 
		 summarise(paper2 = sum(paper, na.rm = TRUE), n = n()) 
  df2$paper <- df2$paper2; df2$paper2 <- NULL # Rename
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

  p <- ggplot(df2, aes(y = paper,
                       x = publication_decade,
  	       	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
             aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	             shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(catal)
  print(p)
}
```

### 2016-07-18 inst/examples/manuscript-krakow.Rmd
Rmd update

```{r topics, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width="430px", fig.show="hold"}
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
df = df0

  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

  # Selected catalogue with selected topics
  df <- df %>% filter(catalog == catalogue & hit)
  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }

@@

lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}
dfl$language <- factor(dfl$language, levels = langs)

>>>>

```{r topics, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, out.width="430px", fig.show="hold"}
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
df <- df0
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)

  # Selected catalogue with selected topics
  df <- df %>% filter(catalog == catalogue & hit)
  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    if (nrow(dflsub)>0) {
      dflsub$language <- gsub("language.", "", lan)
      dfl <- bind_rows(dfl, dflsub)
    }
  }

@@

lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
dfl$language <- factor(dfl$language, levels = langs)
